Machine Learning Autoencoder Applied to Communication Channels

### 2019-06-16 Rapport/Article/Article.tex
Seance 06/16/2019

However, the MAP decoder is practically nonviable to implement \cite{journals/ett/RobertsonHV97}. Thus, its variants cited in this paragraph are better suited for practical cases. The Log-MAP is an optimal decoder, having equivalent performance to the MAP decoder. While the Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.  these variants are widely used as decoders for turbo codes \cite{b7}. Refer to \cite{journals/ett/RobertsonHV97} for specific details in the derivation of the Log-MAP algorithm. Furthermore, the Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is a bit-wise MAP rule with lower complexity widely used in practice, even for long data packets.

Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a BSC was studied in \cite{b5}. They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{journals/ett/RobertsonHV97}. It means that if the real SNR is smaller, the decoder will not perform well. Hence, for channel characteristics that change over time, the application of a MAP algorithm is debatable. This conclusion serve as motivation for ML autoencoders, which could adapt to different SNR conditions, outperforming the MAP decoder for realistic applications.

@@

% TODO: WRITE WHICH ACTIVATION FUNCTION I WILL USE
%
Two common activation functions are the rectified linear unit (ReLU) and the sigmoid function. Both will be applied in the NN treated in this paper. Their mathematical expressions are shown below.

\begin{align}
	\text{ReLU: } \sigma_1 (x)& = max\{0, x\} \label{eq:relu}\\ 
	\text{Sigmoid: } \sigma_2 (x)& = \frac{1}{1+e^{- \lambda x}} \label{eq:sig}
\end{align}
where $x \in \mathbb{R}$ and $\lambda>0$.

@@

One of the greatest advantage of NN is its ability to generalize the training set, delivering the right results even for input data not contained in the training set. With the weight matrix trained, the NN finds the correct output values for unseen inputs. In order to validate this principle, a \textit{validation data set} is used and the loss is mesured. Based on this value, we can evaluate the NN overall accuracy.

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/20190607-160812_MAP}
    \caption{MAP algorithm performance in terms of BER versus channel crossover probability.}\label{fig:MAP}
\end{figure}

@@

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Decoder} & Dense: X, activation: ReLU, input size: $n$ \\
 & Dense: X, activation: ReLU  \\
 & Dense: $k$, activation: Sigmoid  \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
    \thickhline
  \end{tabular}
    \label{tab:arraydecoder}
\end{center}
\end{table}

@@

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Decoder} & Dense: X, activation: ReLU, input size: $n$ \\
 & Dense: X, activation: ReLU  \\
 & Dense: $2^k$, activation: Sigmoid  \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: }XX}\\
    \thickhline
  \end{tabular}
    \label{tab:onehotdecoder}
\end{center}
\end{table}

@@

\section{Results and Analysis}

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MAP_MLNN1H_p_0}
    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/20190529-142119_MAP_MLNN_Mep_65536}
    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
\end{figure}

@@

@@

\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}

\bibliographystyle{ieeetr}
\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}

>>>>

However, the MAP decoder is practically nonviable to implement \cite{journals/ett/RobertsonHV97}. Thus, its variants cited in this paragraph are better suited for practical cases. The Log-MAP is an optimal decoder, having equivalent performance to the MAP decoder. While the Max-Log-MAP is a suboptimal decoder and will not be treated in this paper.  these variants are widely used as decoders for turbo codes \cite{HagenauerJ}. Refer to \cite{journals/ett/RobertsonHV97} for specific details in the derivation of the Log-MAP algorithm. Furthermore, the Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is a bit-wise MAP rule with lower complexity widely used in practice, even for long data packets.

Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a BSC was studied in \cite{JordanMA}. They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{journals/ett/RobertsonHV97}. It means that if the real SNR is smaller, the decoder will not perform well. Hence, for channel characteristics that change over time, the application of a MAP algorithm is debatable. This conclusion serve as motivation for ML autoencoders, which could adapt to different SNR conditions, outperforming the MAP decoder for realistic applications.

@@

% TODO: WRITE WHICH ACTIVATION FUNCTION I WILL USE
%
Two common activation functions are the rectified linear unit (ReLU) and the sigmoid function. Both will be applied to the DNNs treated in this paper. In addition, the Softmax activation function is commonly used in multi classification problems and will be used to the One-hot decoding DNN architectures. Their mathematical expressions are shown below.

\begin{align}
	\text{ReLU: } \sigma_1 (x)& = max\{0, x\} \label{eq:relu}\\ 
	\text{Sigmoid: } \sigma_2 (x)& = \frac{1}{1+e^{- \lambda x}} \label{eq:sig}\\
	\text{Softmax: } \sigma_3 (\textbf{z})& = \frac{e^{z_i}}{\sum_{j=1}^{K}e^{z_j}}
\end{align}
where $x \in \mathbb{R}$, $\lambda>0$, $i = 1,...,K$ and $\textbf{z} = (z_1,...,z_K) \in \mathbb{R}^K$.

@@

One of the greatest advantage of NN is its ability to generalize the training set, delivering the right results even for input data not contained in the training set. With the weight matrix trained, the NN finds the correct output values for unseen inputs. In order to validate this principle, a \textit{validation data set} is used and the loss is measured. Based on this value, we can evaluate the NN overall accuracy.

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MAP-alone-curve}
    \caption{MAP algorithm performance in terms of BER versus channel crossover probability.}\label{fig:MAP}
\end{figure}

@@

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Decoder} & Dense: 128, activation: ReLU, input size: $n$ \\
 & Dense: 64, activation: ReLU  \\
 & Dense: 32, activation: ReLU  \\
 & Dense: $k$, activation: Sigmoid  \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: $12776$}}\\
    \thickhline
  \end{tabular}
    \label{tab:arraydecoder}
\end{center}
\end{table}

@@

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{1}{*}{Decoder} & Dense: 256, activation: Softmax, input size: $n$ \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: }$4352$}\\
    \thickhline
  \end{tabular}
    \label{tab:onehotdecoder}
\end{center}
\end{table}

@@

\subsection{Monte Carlo Simulation}
Explain the prediction simulation (global reps, values of p, etc)
\section{Results and Analysis}

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MLNN_Mep_65536_ptrain_007}
    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
\end{figure}

@@

\begin{table}
\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the mean MAP algorithm decoding time. } 
\begin{center}

  \begin{tabular}{ l | l | l | l | l }
    \thickhline
    MAP & A. Dec. & One-hot Dec. & A. Auto. & One-hot Auto. \\ \hline
    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ & $ \pm $ & $ \pm $ \\ 
    \thickhline
  \end{tabular}
    \label{tab:timeanalysis}
\end{center}
\end{table}

@@

\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}

\bibliographystyle{ieeetr}
\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18}

### 2019-06-16 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/16/2019

\subsection{Design \& Architecture}
\begin{frame}{Neural Network's Architecture}{Show the architecture used for each case and remarks some important parameters}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

@@

\section{Results \& Discussions}
\subsection{DNN Decoders}
\begin{frame}{DNN Array Decoder}{Show the results for the array decoder in terms of train p, Mep, Parameters, etc}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

\begin{frame}{DNN One-hot Decoder}{Show the results for the one-hot decoder in terms of train p, Mep, Parameters, etc}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

@@

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{Bibliography}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{Bibliography}

  \cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}

  \bibliographystyle{ieeetr}
  \bibliography{../../Article/bib/ShannonCE1948,../../Article/bib/CalabreseWGPS16,../../Article/bib/OSheaH17,../../Article/bib/Oshea2,../../Article/bib/Gruber,../../Article/bib/Worm,../../Article/bib/Viterbi,../../Article/bib/RobertsonHV97,../../Article/bib/Ibnkahla,../../Article/bib/Nielsen,../../Article/bib/Murphy,../../Article/bib/AbadiABBCCCDDDG16,../../Article/bib/Keras,../../Article/bib/mit_neco18_1527,../../Article/bib/BenammarP18}

>>>>

\subsection{Design \& Architecture}
\begin{frame}{Neural Network's Architecture}{Show the architecture used for each case and remarks some important parameters}
  \begin{itemize}
  \item {

  }
  \item {

  }
  \item {

  }
  \end{itemize}
\end{frame}

@@

\section{Results \& Discussions}
\subsection{DNN Decoders}
\begin{frame}{DNN Array Decoder}{Show the results for the array decoder in terms of train p, Mep, Parameters, etc}

    \begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{../../Article/images/MLNN_Mep_65536_ptrain_007}
    \caption{Array decoding BER performance. NN trained with a channel crossover probability error of $p_t=0.07$.}\label{fig:ArrayD}
\end{figure}

\begin{frame}{DNN One-hot Decoder}{Show the results for the one-hot decoder in terms of train p, Mep, Parameters, etc}

    \begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{../../Article/images/MLNN1H_Mep_16384_ptrain_0}
    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$.}\label{fig:1HD}
\end{figure}

@@

% All of the following is optional and typically not needed. 
%\appendix
%\section<presentation>*{Bibliography}

%\begin{frame}[allowframebreaks]
%  \frametitle<presentation>{Bibliography}

 % \cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18}

 % \bibliographystyle{ieeetr}
 % \bibliography{../../Article/bib/ShannonCE1948,../../Article/bib/CalabreseWGPS16,../../Article/bib/OSheaH17,../../Article/bib/Oshea2,../../Article/bib/Gruber,../../Article/bib/Worm,../../Article/bib/Viterbi,../../Article/bib/RobertsonHV97,../../Article/bib/Ibnkahla,../../Article/bib/Nielsen,../../Article/bib/Murphy,../../Article/bib/AbadiABBCCCDDDG16,../../Article/bib/Keras,../../Article/bib/mit_neco18_1527,../../Article/bib/BenammarP18}

%\end{frame}
